---
title: "Makeup"
author: "Nhi Hin"
date: "2021-03-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
library(dplyr)
library(magrittr)
library(readr)
library(reshape2)
library(ggplot2)
library(grDevices)
library(plotly)
ggplot2::theme_set(theme_minimal())
```


## Links

- Skin complexion makeup analysis (pudding.cool): https://pudding.cool/2021/03/foundation-names/

- TidyTuesday page: https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-03-30

- Data collection and preprocessing details: https://github.com/the-pudding/data/tree/master/foundation-names

## Import Data

```{r}
sephora <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/sephora.csv')
ulta <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/ulta.csv')
allCategories <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allCategories.csv')
allShades <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allShades.csv')
allNumbers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allNumbers.csv')

```

- The `allCategories` dataset looks like this. It contains information scraped from both Ulta and Sephora. 

```{r}
allCategories %>% dplyr::select(brand, name, categories, hex, lightness) 
```

- How many brands are represented in this dataset?

```{r}
brands <- allCategories %>% 
  dplyr::group_by(brand) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(desc(n))

brands

dim(brands)

```

- There are `r nrow(brands)` represented in this dataset. Multiple products are available for each brand. 

- In terms of categories, there are also many different categories including `food`, `gem`, etc. Because of how substantial the dataset is, it might be easier if we focus on just one category for now. 

```{r}
foodSubset <- allCategories %>% dplyr::filter(categories == "food")
dim(foodSubset)

foodSubset %>% dplyr::select(brand, name, hex, lightness)
```

- Let's have a look at the most common food descriptors through counting the `name` column.

```{r}
foodSubset$name %>% 
  table %>% 
  as.data.frame %>% 
  set_colnames(c("name", "n")) %>%
  dplyr::arrange(desc(n))
```

- The descriptors `bisque`, `honey`, `almond`, `chestnut`, and `vanilla` appear the most common. Is this due to the same brands, or does it seem to be common across all brands?

```{r}
foodSubset_summary <- foodSubset %>%
  dplyr::filter(name %in% c("bisque",
                            "honey",
                            "almond", 
                            "chestnut",
                            "vanilla"))  %>% 
  dplyr::group_by(brand, name) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(name, desc(n)) 

foodSubset_summary
```

## Variation in Makeup Shades Named After Food

- Let's create a quick stacked bar chart to show the relative proportions of each brand within each of the top 5 food descriptors. 

```{r}
# Get a large number of distinct colours, see
# https://stackoverflow.com/questions/15282580/how-to-generate-a-number-of-most-distinctive-colors-in-r
library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

foodSubset_summary %>%
  ggplot(aes(x = name, y = n, fill = brand)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = col_vector) +
  labs(y = "Number", x = "Food Name", fill = "Brand") +
  ggtitle("Most popular food descriptors in foundation shades", 
          subtitle = "Brand composition") 
```

- It might be more informative if we can actually use the Hex color code provided to show any differences / similarities in the different makeup products. 

```{r fig.width=11}
test <- foodSubset %>%
  dplyr::filter(name %in% c("bisque",
                            "honey",
                            "almond", 
                            "chestnut",
                            "vanilla",
                            "caramel"))  %>% 
  split(., f=.$name) %>%
  lapply(function(x){
    x %>% dplyr::arrange(desc(as.numeric(lightness)))
  }) %>%
  dplyr::bind_rows(.id = "name") %>%
  ggplot(aes(x = n, fill = factor(hex, levels = unique(hex)))) +
  geom_bar(stat = "count") +
  scale_fill_identity() +
  facet_wrap(~factor(name, levels = c("chestnut",
                                      "caramel", 
                                      "almond",
                                      "honey", 
                                      "bisque", 
                                      "vanilla")), nrow = 1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
 
test
```

## Shade Similarity Exploration

- An interesting idea I had was whether it's possible to pick a makeup shade similar to another one. In other words, if we can describe the different makeup shades in terms of colour code similarity, it might make it easier to see which makeup shades are essentially the same from different brands. 

### Conversion of HEX Colour Code to RGB

- It is possible to measure the distance between colours (Euclidean distance) from RBG but as far as I'm aware, it isn't as straightforward with Hex colour codes, which is what we've been given in the `allShades` data. 

- First, we will convert Hex to RGB using the `col2rgb` function. The following will add 3 columns to `allShades` corresponding to Red, Green, and Blue. 

```{r}
allShades <- allShades %>%
  rowwise() %>%
  dplyr::mutate(red = col2rgb(hex)[1],
                green = col2rgb(hex)[2],
                blue = col2rgb(hex)[3]) 

```

- Let's see if we can use the RGB values to do a PCA. 

```{r}

pcAnalysis <- allShades %>% 
  dplyr::select(red, green, blue) %>% 
  prcomp()

summary(pcAnalysis)

# PC1 and PC2
pcPlot <- pcAnalysis$x %>%
  as.data.frame %>%
  dplyr::mutate(hex = allShades$hex) %>%
  dplyr::filter(PC2 < 150) %>% # Remove outlier sample PC2=181 which comes up as colour blue
  ggplot(aes(x = PC1, y = PC2, color = hex)) +
  geom_point()+
  scale_color_identity() +
  labs(x = "PC1 (95%)", y = "PC2 (4.4%)") +
  ggtitle("Principal Component Analysis of All Shades") 

pcPlot

# PC1 and PC3
pcPlot2 <- pcAnalysis$x %>%
  as.data.frame %>%
  dplyr::mutate(hex = allShades$hex) %>%
  dplyr::filter(PC2 < 150) %>%
  ggplot(aes(x = PC1, y = PC3, color = hex)) +
  geom_point()+
  scale_color_identity() +
  labs(x = "PC1 (95%)", y = "PC3 (0.3%)") +
  ggtitle("Principal Component Analysis of All Shades") 

pcPlot2

# 3D plot - not run
# pca3d::pca3d(pcAnalysis, col = allShades$hex)
```

## Using Plotly

```{r}
dataForPlotting <- allShades %>% 
  dplyr::select(red, green, blue) %>% 
  bind_cols(pcAnalysis$x %>% as.data.frame) 
  dplyr::filter(PC2 < 150)

fig <- plot_ly(data = dataForPlotting,
               x = ~PC1, y = ~PC2, z = ~PC3, 
               color = ~hex, colors = ~hex) %>% 
  layout(showlegend = FALSE)

fig
```


