---
title: "Makeup"
author: "Nhi Hin"
date: "2021-03-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
library(dplyr)
library(magrittr)
library(readr)
library(reshape2)
library(ggplot2)
ggplot2::theme_set(theme_minimal())
```


## Links

- Skin complexion makeup analysis (pudding.cool): https://pudding.cool/2021/03/foundation-names/

- TidyTuesday page: https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-03-30

- Data collection and preprocessing details: https://github.com/the-pudding/data/tree/master/foundation-names

## Import Data

```{r}
sephora <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/sephora.csv')
ulta <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/ulta.csv')
allCategories <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allCategories.csv')
allShades <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allShades.csv')
allNumbers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allNumbers.csv')

```

- The `allCategories` dataset looks like this. It contains information scraped from both Ulta and Sephora. 

```{r}
allCategories %>% dplyr::select(brand, name, categories, hex, lightness) 
```

- How many brands are represented in this dataset?

```{r}
brands <- allCategories %>% 
  dplyr::group_by(brand) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(desc(n))

brands

dim(brands)

```

- There are `r nrow(brands)` represented in this dataset. Multiple products are available for each brand. 

- In terms of categories, there are also many different categories including `food`, `gem`, etc. Because of how substantial the dataset is, it might be easier if we focus on just one category for now. 

```{r}
foodSubset <- allCategories %>% dplyr::filter(categories == "food")
dim(foodSubset)

foodSubset %>% dplyr::select(brand, name, hex, lightness)
```

- Let's have a look at the most common food descriptors through counting the `name` column.

```{r}
foodSubset$name %>% 
  table %>% 
  as.data.frame %>% 
  set_colnames(c("name", "n")) %>%
  dplyr::arrange(desc(n))
```

- The descriptors `bisque`, `honey`, `almond`, `chestnut`, and `vanilla` appear the most common. Is this due to the same brands, or does it seem to be common across all brands?

```{r}
foodSubset_summary <- foodSubset %>%
  dplyr::filter(name %in% c("bisque",
                            "honey",
                            "almond", 
                            "chestnut",
                            "vanilla"))  %>% 
  dplyr::group_by(brand, name) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(name, desc(n)) 

foodSubset_summary
```

- Let's create a quick stacked bar chart to show the relative proportions of each brand within each of the top 5 food descriptors. 

```{r}
# Get a large number of distinct colours, see
# https://stackoverflow.com/questions/15282580/how-to-generate-a-number-of-most-distinctive-colors-in-r
library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

foodSubset_summary %>%
  ggplot(aes(x = name, y = n, fill = brand)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = col_vector) +
  labs(y = "Number", x = "Food Name", fill = "Brand") +
  ggtitle("Most popular food descriptors in foundation shades", 
          subtitle = "Brand composition") 
```

- It might be more informative if we can actually use the Hex color code provided to show any differences / similarities in the different makeup products. 

```{r fig.width=11}
test <- foodSubset %>%
  dplyr::filter(name %in% c("bisque",
                            "honey",
                            "almond", 
                            "chestnut",
                            "vanilla",
                            "caramel"))  %>% 
  split(., f=.$name) %>%
  lapply(function(x){
    x %>% dplyr::arrange(desc(as.numeric(lightness)))
  }) %>%
  dplyr::bind_rows(.id = "name") %>%
  ggplot(aes(x = n, fill = factor(hex, levels = unique(hex)))) +
  geom_bar(stat = "count") +
  scale_fill_identity() +
  facet_wrap(~factor(name, levels = c("chestnut",
                                      "caramel", 
                                      "almond",
                                      "honey", 
                                      "bisque", 
                                      "vanilla")), nrow = 1)  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
 
test
```

