---
title: "Plastics"
author: "ofchurches"
date: "2 February 2021"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r include=FALSE, warning=FALSE}
library(readr)
library(ggalt)
library(magrittr)
library(stringr)
library(ggplot2)
library(dplyr)
library(reshape2)
theme_set(theme_bw())
```


# Background

The read.me with instructions, notes, data and data-dictionary is [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-26/readme.md).

# Get the data

```{r message=FALSE, warning=FALSE}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')
```

## Clean up data

- Remove obvious typos/mistakes:

```{r}
clean_plastics <- plastics %>%
  dplyr::mutate(parent_company = str_to_title(parent_company),
                country = str_to_title(country)) %>% 
  dplyr::filter(!parent_company %in% c("#Error!", "Unbranded", "null",
                                       "NULL","Null", "Grand Total"),
                !country %in% c("EMPTY"))
```

- Remove duplicated rows

```{r}
clean_plastics2<- clean_plastics %>% 
  split(x = ., f=.$country) %>%
  lapply(function(x){
    x %>% dplyr::arrange(parent_company, desc(grand_total)) %>%
  dplyr::distinct(country, year, parent_company, num_events, volunteers, .keep_all=TRUE)
  }) %>% bind_rows()
```

- Replace NAs for plastic collected with 0:

```{r}
# Get the columns associated with plastics collected 
# and replace NA with 0
clean_plastics3 <- clean_plastics2[, c(4:12)]
clean_plastics3[is.na(clean_plastics3)] <- 0

clean_plastics2[, c(4:12)] <- clean_plastics3
```


## Initial data exploration

### Countries' participation in 2019 and 2020

```{r}
countries_by_year <- clean_plastics2 %>% 
  dplyr::group_by(country, year) %>% 
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(country, year)

# 50 countries participated in 2019
countries_2019 <- countries_by_year %>% 
  dplyr::filter(year == 2019) %>%
  as_tibble
countries_2019

# 55 countries participated in 2020
countries_2020 <- countries_by_year %>% 
  dplyr::filter(year == 2020) %>%
  as_tibble 
countries_2020

# 38 countries participated in both years
inner_join(countries_2019,
           countries_2020, 
           by = c("country")) %>% 
  as_tibble()
```

### How many parent companies participated per country

```{r}
companies_by_country_year <- clean_plastics2 %>%
  dplyr::select(country, parent_company) %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(n_companies = n()) %>%
  dplyr::arrange(desc(n_companies))
companies_by_country_year 
```

## Visualising the types of plastics collected per country in 2020

```{r}
grandtotals <- clean_plastics2 %>%
  dplyr::filter(year == 2020) %>%
  dplyr::select(country, grand_total, volunteers) %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(sum = sum(grand_total),
                   sum_volunteers = sum(volunteers))  %>%
  dplyr::mutate(norm_total = sum/sum_volunteers)

clean_plastics2 %>% 
  dplyr::filter(year == 2020) %>%
  dplyr::select(country, parent_company, hdpe, ldpe, o, pet, pp, ps, pvc) %>%
  melt(measure.vars= c("hdpe", "ldpe", "o", "pet", "pp", "ps", "pvc")) %>% 
  dplyr::rename(plasticType = variable) %>%
  dplyr::left_join(grandtotals, by = "country") %>%
  ggplot(aes(x = reorder(country, sum), y = value, fill = plasticType)) +
  geom_bar(stat="identity") +
  coord_flip() +
  scale_fill_brewer(palette="Accent") +
  labs(x = "Number of units", y = "Country")+
  ggtitle("Plastics collected in 2020")
```

- And 2019:

```{r}
grandtotals <- clean_plastics2 %>%
  dplyr::filter(year == 2019) %>%
  dplyr::select(country, grand_total) %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(sum = sum(grand_total))

clean_plastics2 %>% 
  dplyr::filter(year == 2019) %>%
  dplyr::select(country, parent_company, hdpe, ldpe, o, pet, pp, ps, pvc) %>%
  melt(measure.vars= c("hdpe", "ldpe", "o", "pet", "pp", "ps", "pvc")) %>% 
  dplyr::rename(plasticType = variable) %>%
  dplyr::left_join(grandtotals, by = "country") %>%
  ggplot(aes(x = reorder(country, sum), y = value, fill = plasticType)) +
  geom_bar(stat="identity") +
  coord_flip() +
  scale_fill_brewer(palette="Accent") +
  labs(x = "Number of units", y = "Country") +
  ggtitle("Plastics collected in 2019")
```


